Beta Diversity through Time and Space

### 2020-09-15 analyses/beta_diversity.Rmd
refined GDM analysis

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(vegan)
library(lubridate)
library(gdm)
library(bipartite)
library(sf)
library(mgcv)
library(mgcViz)
library(tidyverse)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv")%>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

site.year_transects <- net %>%
  group_by(site, year) %>%
  summarize(transects = length(unique(date)))

site_transects <- net %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

web_site.year <- net %>%
  group_by(site.year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2010 <- net %>%
  filter(year == 2010) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2011 <- net %>%
  filter(year == 2011) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2012 <- net %>%
  filter(year == 2012) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)
array_site.date.2010 <- webs2array(web_site.date.2010)
array_site.date.2011 <- webs2array(web_site.date.2011)
array_site.date.2012 <- webs2array(web_site.date.2012)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

#write_csv(beta_site, "../analyses/output/beta_site.csv")

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

### By individual transect

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "date1"), sep = "_") %>%
  separate(j, into = c("site2", "date2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)

beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE)
```

# Beta diversity by elevation difference
```{r}
# Interaction beta diversity by elevation difference (by year)
## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1)

ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1, scales = "free_x")

# Interaction beta diversity by elevation difference, this time pooling across years
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

# How important is the difference in transects? Not as important as the difference in elevation.
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(beta_site, 
       aes(elev.diff, transects.diff)) +
  geom_point(alpha = 0.05) +
  geom_smooth()

### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

ST_diff <- beta_site %>%
  filter(metric == "ST") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

OS_diff <- beta_site %>%
  filter(metric == "OS") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

elev_diff <- beta_site %>%
  dplyr::select(site1, site2, elev.diff) %>%
  unique() %>%
  spread(site2, elev.diff) %>%
  dist()

transects_diff <- beta_site %>%
  dplyr::select(site1, site2, transects.diff) %>%
  unique() %>%
  spread(site2, transects.diff) %>%
  dist()

WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
```

### Let's look at floral survey diversity patterns
```{r}
survey_transects <- survey %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

survey_bioData <- survey %>%
  filter(flower.cover > 0) %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()
%>%
  left_join(survey_transects)

survey_sitepair <- formatsitepair(survey_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_survey <- gdm(survey_sitepair, geo = TRUE)
gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey)
plot(gdm_survey)
plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### And now Bombus diversity patterns
```{r}
BB_bioData <- net %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count)) %>%
  group_by(site, bb.sp) %>%
  summarize(abund = sum(prop)) %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()

bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_bb <- gdm(bb_sitepair, geo = TRUE)
gdm.varImp(bb_sitepair, geo = TRUE)

summary(gdm_bb)
plot(gdm_bb)
plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### Now let's look at interaction beta diversity
```{r}
net_sitepair <- beta_site %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(net_sitepair, geo = TRUE)
gdm.varImp(net_sitepair, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

>>>>

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(lubridate)
library(sf)
library(gdm)
library(bipartite)
library(vegan)
library(mgcv)
library(mgcViz)
library(see)
library(tidyverse)
```

### Climate data
```{r echo=FALSE, message=FALSE, include=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

### Load network and survey data
```{r}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% # create site.date field that can be split up again later
  mutate(year = factor(year)) # convert year to factor

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
# Function to convert networks to webs arrays
web_prep <- function(net, y) {
  out <- net %>%
    filter(year == y) %>%
    unite(site.date, c(site, date), sep = "_") %>%
  group_by(site.date, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  group_by(site.date) %>%
  mutate(bb.sp.count = length(unique(bb.sp)),
         plant.sp.count = length(unique(plant.sp))) %>%
  ungroup() %>%
  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% # the betalinkr function fails if you feed it webs with only 1 species in either (or perhaps both) of the levels; this is probably for good reason.
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs(type.out = "array")
}

# Create webs
net2010 <- web_prep(net, 2010)
net2011 <- web_prep(net, 2011)
net2012 <- web_prep(net, 2012)

# Function with betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}

# Calculate beta-diversity
beta_site.date.2010 <- betalinkr_set1(net2010)
beta_site.date.2011 <- betalinkr_set1(net2011) 
beta_site.date.2012 <- betalinkr_set1(net2012)

# Collate results into a single tibble
beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)

# Save to file to avoid redoing that long step
write_csv(beta_site.date, "./output/beta_site.date.csv")
beta_site.date <- read_csv("./output/beta_site.date.csv")

# Add site data
beta_mit_data <- beta_site.date %>%
  separate(i, into = c("site1", "date1"), sep = "_") %>%
  separate(j, into = c("site2", "date2"), sep = "_") %>%
  mutate(date1 = ymd(date1),
         date2 = ymd(date2),
         year1 = factor(year(date1)),
         year2 = factor(year(date2)),
         yday1 = yday(date1),
         yday2 = yday(date2)) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, man1 = management, x1 = x, y1 = y) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, man1, man2 = management, x1, y1, x2 = x, y2 = y) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           x1, y1, x2, y2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich"))
    )
```

### Plot interaction beta-diversity against elevation difference and time difference
```{r}
# Set palette
p <- c("OS" = "darkorange1", "ST" = "green3", 
       "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
       "ST.lh" = "deepskyblue", "WN" = "black")

# Elevation diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition")), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_manual(values = p) +
  theme_abyss()

# Time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition")), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_manual(values = p) +
  theme_abyss()

# Elevation diff within 1 wk time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition") &
                yday.diff <= 7), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.1, pch = 16, size = 1) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_manual(values = p, name = "partition") +
  xlab("Elevation difference") +
  ylab("β-diversity")
ggsave("./output/beta_elev.png", height = 4.5, width = 6.5, units = "in")

# Time diff within 250 m elev diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition") &
                elev.diff <= 250), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.1, pch = 16, size = 1) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_manual(values = p, name = "partition") +
  xlab("Elevation difference") +
  ylab("β-diversity")
ggsave("./output/beta_time.png", height = 4.5, width = 6.5, units = "in")
```

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Site_pair prep function
site_pair_prep_net <- function(dat, y) {
  beta_mit_data %>%
    ungroup() %>%
    filter(year1 == year2 & 
             year1 == y & 
             site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
    filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
    mutate(weights = rep(1, n())) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
    dplyr::select(distance = value, weights,
                  s1.xCoord = x1, s1.yCoord = y1, 
                  s2.xCoord = x2, s2.yCoord = y2, 
                  s1.elev.mean = elev1, s1.yday = yday1,
                  s2.elev.mean = elev2, s2.yday = yday2) %>%
    data.frame()
}

# Prep site-pairs for GDM analysis
sitepair_2010.WN <- site_pair_prep_net(beta_mit_data, 2010)
sitepair_2011.WN <- site_pair_prep_net(beta_mit_data, 2011)
sitepair_2012.WN <- site_pair_prep_net(beta_mit_data, 2012)

# Run GDM analysis
gdm_WN.2010 <- gdm(sitepair_2010.WN, geo = TRUE)
gdm_WN.2011 <- gdm(sitepair_2011.WN, geo = TRUE)
gdm_WN.2012 <- gdm(sitepair_2012.WN, geo = TRUE)

summary(gdm_WN.2010)
summary(gdm_WN.2011)
summary(gdm_WN.2012)

# Permutation tests (slow)
gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE)
gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE)

# Plot
plot(gdm_WN.2010, include.rug = TRUE, 
     rug.sitepair = sitepair_2010.WN, 
     plot.layout = c(2,3))
plot(gdm_WN.2011, include.rug = TRUE, 
     rug.sitepair = sitepair_2011.WN, 
     plot.layout = c(2,3))
plot(gdm_WN.2012, include.rug = TRUE, 
     rug.sitepair = sitepair_2012.WN, 
     plot.layout = c(2,3))

# Plot with confidence intervals (slow; also, I'm not sure about the current settings)
plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
plotUncertainty(sitepair_2012.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)

# Plot nicely
## wrapper for extracting splines and prepping tibble for plotting
spline_prep <- function(x, y) {
  isplineExtract(x) %>%
    as.data.frame() %>%
    tibble() %>%
    mutate(x.Geographic = x.Geographic/1000) %>%
    pivot_longer(cols = c(x.Geographic, x.elev.mean, x.yday), 
                 names_to = "x.var", values_to = "predictor", 
                 names_prefix = "x.") %>%
    pivot_longer(cols = c(y.Geographic, y.elev.mean, y.yday), 
                 names_to = "y.var", values_to = "response", 
                 names_prefix = "y.") %>%
    filter(x.var == y.var) %>%
    mutate() %>%
    dplyr::select(-y.var) %>%
    mutate(year = factor(rep(y, n())))
}

## extract splines
interaction_splines2010 <- spline_prep(gdm_WN.2010, 2010)
interaction_splines2011 <- spline_prep(gdm_WN.2011, 2011)
interaction_splines2012 <- spline_prep(gdm_WN.2012, 2012)

## collate extracted splines  
interaction_splines <- bind_rows(interaction_splines2010, 
                                 interaction_splines2011, 
                                 interaction_splines2012) %>%
  mutate(x.var = case_when(
    x.var == "elev.mean" ~ "Elevation (m)",
    x.var == "Geographic" ~ "Geographic (km)",
    x.var == "yday" ~ "Day of Year"),
    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
    )

## plot
ggplot(interaction_splines, aes(predictor, response)) +
  geom_line() +
  facet_grid(year ~ x.var, scales = "free_x") +
  xlab(NULL) +
  ylab("f(x)") +
  theme_gray(12)
ggsave("./output/interaction_GDM.png")
```

### GDM: Floral survey
```{r}
# Create functions for prepping data
biodata_prep_survey <- function(surv, y) {
  surv %>%
    mutate(year = year(date),
           yday = yday(date)) %>%
    filter(flower.cover > 0 & year == y) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, yday, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site, yday) %>%
    unite(site.day, site, yday, sep = "_")
}

predData_prep_survey <- function(surv, y) {
  surv %>%
    mutate(year = year(date),
           yday = yday(date)) %>%
    filter(flower.cover > 0 & year == y) %>%
    dplyr::select(site, yday) %>%
    left_join(site_data) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, y, x, yday) %>%
    unique() %>%
    arrange(site, yday) %>%
    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
    dplyr::select(-site)
}

format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site.day", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE)
}

# Prep input for GDM
survey_bioData.2010 <- biodata_prep_survey(survey, 2010)
survey_predData.2010 <- predData_prep_survey(survey, 2010)
survey_sitepair.2010 <- format_set(survey_bioData.2010, survey_predData.2010)

survey_bioData.2011 <- biodata_prep_survey(survey, 2011)
survey_predData.2011 <- predData_prep_survey(survey, 2011)
survey_sitepair.2011 <- format_set(survey_bioData.2011, survey_predData.2011)

survey_bioData.2012 <- biodata_prep_survey(survey, 2012)
survey_predData.2012 <- predData_prep_survey(survey, 2012)
survey_sitepair.2012 <- format_set(survey_bioData.2012, survey_predData.2012)

# Call GDMs
gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)

summary(gdm_survey.2010)
summary(gdm_survey.2011)
summary(gdm_survey.2012)

# Permutations tests (slow)
gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE)

# Plot
plot(gdm_survey.2010)
plot(gdm_survey.2011)
plot(gdm_survey.2012)

# Plot with confidence intervals (slow)
plotUncertainty(survey_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
plotUncertainty(survey_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)

# Plot nicely
## extract splines
survey_splines2010 <- spline_prep(gdm_survey.2010, 2010)
survey_splines2011 <- spline_prep(gdm_survey.2011, 2011)
survey_splines2012 <- spline_prep(gdm_survey.2012, 2012)

## collate extracted splines  
survey_splines <- bind_rows(survey_splines2010,
                            survey_splines2011, 
                            survey_splines2012) %>%
  mutate(x.var = case_when(
    x.var == "elev.mean" ~ "Elevation (m)",
    x.var == "Geographic" ~ "Geographic (km)",
    x.var == "yday" ~ "Day of Year"),
    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
    )

## plot
ggplot(survey_splines, aes(predictor, response)) +
  geom_line() +
  facet_grid(year ~ x.var, scales = "free_x") +
  xlab(NULL) +
  ylab("f(x)") +
  theme_gray(12)
ggsave("./output/survey_GDM.png")
```
